---
title: "VisumAnalysis"
author: "Dario Righelli"
date: "2/21/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Installation

Installing and loading VisualExperiment class.

```{r}
library(devtools)
install_github("drighelli/SpatialExperiment", ref="spatialExperiment")
```


```{r}
library("SpatialExperiment")
```


# Data Loading

Examples data files have been downloaded from the 10x site, [Mouse adult brain](https://support.10xgenomics.com/spatial-gene-expression/datasets/1.0.0/V1_Adult_Mouse_Brain)

## Features

Loading features and count matrix.

```{r}
barcodes <- read.csv("data/visium/mouse_coronal_V1/raw_feature_bc_matrix/barcodes.tsv", sep="\t", 
                     header=FALSE, col.names=c("Barcodes"))

barcodesEx <- head(barcodes, n=50)
saveRDS(object=barcodesEx, file="barcodesEx.Rds")
features <- read.csv("data/visium/mouse_coronal_V1/raw_feature_bc_matrix/features.tsv", sep="\t", 
                     header=FALSE, col.names=c("Barcodes", "Feature_name", 
                                               "Feature_type"))
featuresEx <- head(features, n=50)
saveRDS(object=featuresEx, file="featuresEx.Rds")
library("Matrix")
counts <- readMM(file="data/visium/mouse_coronal_V1/raw_feature_bc_matrix/matrix.mtx")
countsEx <- head(counts, n=50)[,1:50]
saveRDS(object=countsEx, file="countsEX.Rds")
```

## Spatial Coordinates

Loading Spatial coordinates of the tissue section.

```{r}


tissue.positions <- read.csv("data/visium/mouse_coronal_V1/spatial/tissue_positions_list.csv", 
                             header=FALSE, col.names=c("Barcodes", "in_tissue", 
                                                       "array_row", "array_col",
                                                       "pxl_col_in_fullres", "pxl_row_in_fullres"))
tissPosEx <- tissue.positions[match(barcodesEx$Barcodes, tissue.positions$Barcodes),]
saveRDS(object=tissPosEx, file="tissPosEx.Rds")
```

## Spatial Scale Factors

Scale factors for the given three image formats transformations.

```{r}
library("rjson")
scalefactors <- fromJSON(file="data/visium/mouse_coronal_V1/spatial/scalefactors_json.json")
```


# VisiumExperiment

## Creation

Creating VisiumExperiment class within 10x Visium data and testing the show method.

```{r}
ve <- VisiumExperiment(rowData=features, colData=barcodes, 
                           assays=c(counts=counts), 
                           spatialCoords=tissue.positions,
                           scaleFactors=scalefactors)
ve

ve <- VisiumExperiment(rowData=featuresEx, colData=barcodesEx, 
                           assays=c(counts=countsEx), 
                           spatialCoords=tissPosEx,
                           scaleFactors=scalefactors)
ve
```

## Methods

### Retrieving Spatial Coordinates

The *spatialCoords* methods can be used to retrieve the spatial coordinates., 
It can be used also to update the values.

```{r}
spatialCoords(ve)
```

### Number of array barcoded-spots covered by tissue

The *isInTissue* method can be used to know which and how many spots are covered
by the tissue.

```{r}
sum(isInTissue(ve))
# ve[isInTissue(ve)] ## subset not defined yet
```

