---
title: "SpatialExperimentWorkflow"
author: "Dario Righelli"
date: "3/8/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading data

We stored the provided seqFish and scRNA-seq data inside the data/seqFISH folder

## seqFISH data

```{r}
fishCoordinates <- read.table(file="../SpatialAnalysis/data/seqFISH/fcortex.coordinates.txt", header=FALSE, sep=" ")
colnames(fishCoordinates) <- c("Cell_ID", "Irrelevant", "x", "y")
save(fishCoordinates, file="seqFISH_Coordinates.rda")

fishCellLabels <- read.table(file="../SpatialAnalysis/data/seqFISH/seqfish_cell_labels.tsv", header=FALSE, sep="\t")
colnames(fishCellLabels) <- c("Cell_ID", "cluster", "class", "classID", "Irrelevant", "Prob")
save(fishCellLabels, file="seqFISH/seqFISH_Labels.rda")

fishFeaturesCounts <- read.table(file="../SpatialAnalysis/data/seqFISH/seqfish_normalized_cortex_b2_testing.txt", header=FALSE, sep="\t", row.names=1)
fishFeaturesCounts <- as.matrix(fishFeaturesCounts)
save(fishFeaturesCounts, file="seqFISH/seqFISH_Counts.rda")
```

## Single Cell RNA-seq data

```{r}
scRNASeqCounts <- read.table(file="../SpatialAnalysis/data/seqFISH/tasic_scRNAseq_training_b2.txt", header=FALSE, sep="\t", row.names=1)
scRNASeqCounts <-  as.matrix(scRNASeqCounts)
save(scRNASeqCounts, file="seqFISH/scRNA_Counts.rda")

scRNASeqLabels <- read.table(file="../SpatialAnalysis/data/seqFISH/tasic_scRNAseq_labels.tsv", header=FALSE, sep="\t")
save(scRNASeqLabels, file="seqFISH/scRNA_Labels.rda")
```

# SingleCellExperiment package

We load the scRNA-seq data into a SingleCellExperiment class, so we can use it
for further joint analysis with the seqFISH data.

```{r}
library(SingleCellExperiment)
sce <- SingleCellExperiment(rowData=rownames(scRNASeqCounts), 
                    colData=scRNASeqLabels,
                    assays=SimpleList(counts=as.matrix(scRNASeqCounts)))
show(sce)
```


# SpatialExperiment package

## Installation

We install the SpatialExperiment package which is available through the github/drighelli
repository on spatialExperiment branch.

```{r}
library(devtools)
install_github("drighelli/SpatialExperiment")
```

## Class handling

Once installed, we can load the data inside the SpatialExperiment class.
We then use the show method for looking at the class description.

```{r}
library(SpatialExperiment)
se <- SpatialExperiment(rowData=rownames(fishFeaturesCounts),
                        colData=fishCellLabels,
                        assays=SimpleList(counts=as.matrix(fishFeaturesCounts)),
                        spatialCoords=fishCoordinates)
show(se)

```

## Spatial Coordinates

With the aid of the spatialCoords methods we can get and set the spatial coordinates.

### Getter

```{r}
spatialCoords(se)
```

### Setter

We create a fake fish coordinates data frame and overwrite the old one,
showing that the y coordinates are identical to the x ones.

```{r}
fakeFishCoords <- cbind(fishCoordinates[,c(1:3)], fishCoordinates[,3])
colnames(fakeFishCoords) <- c("Cell_ID", "Irrelevant", "x", "y")
spatialCoords(se) <- fakeFishCoords
spatialCoords(se)
```

